---
title: "Lab1"
author: "Bertrand Wilden"
date: "1/8/2021"
output: pdf_document
---

# This is a big title!
## This one is a little smaller
### Wow even smaller

Typing plain text is easy! Just make sure you are not in a code chunk and go wild! Blah blah blah blah blah.

Make a new paragraph by leaving an empty line between your text.

Lists are easy too! Why I love RMarkdown:

  1. I can reproduce all my analysis in one document
  2. No unecessary frills
  3. Excellent citation integration with Zotero

It even does math. This is an inline equation: $Y_i = \beta_0 + \beta_1X_i + \epsilon_i$. This is a standalone equation:
$$
Y_i = \beta_0 + \beta_1X_i + \epsilon_i
$$

But the real fun begins when we start including R code chunks and output graphs/tables directly in our output pdf document.

```{r message = FALSE}
library(here)
library(readr)
library(ggplot2)
library(dplyr)

lifeexp <- read_csv(here("data", "lifeexp.csv"))

head(lifeexp)
```

This chunk loads in the packages we need and loads the country life expectancy data set. The `head()` command displays the first few rows of the data set so we have an idea about what we're working with. We set `message = FALSE` in the chunk options to prevent the `library()` commands spamming our pdf with extra messages. 

Now lets make some graphs!

## Histograms

```{r warning=FALSE}
# Pipe off from our original data set
lifeexp %>% 
  # Create an empty ggplot object. x = lexp means we will be displaying 
  # the lexp variable
  ggplot(aes(x = lexp)) +
  # geom_* specifies the plot type. Histogram in this case
  geom_histogram()
```

It looks okay but ggplot has a lot more options for customization. Go crazy!

```{r warning=FALSE}
lifeexp %>% 
  ggplot(aes(x = lexp)) +
  geom_histogram(color = "white", fill = "violetred4") + 
  theme_minimal() +
  labs(x = "Life Expectancy (years)", y = "Count", 
       title = "Distribution of Life Expectancy across Countries")
```

## Box Plots

```{r}
lifeexp %>% 
  ggplot(aes(x = lexp)) +
  geom_boxplot(fill = "seagreen") + 
  theme_minimal() +
  labs(x = "Life Expectancy (years)")
```

```{r}
lifeexp %>% 
  ggplot(aes(x = lexp, color = region)) +
  geom_boxplot() + 
  theme_minimal() +
  labs(x = "Life Expectancy (years)")
```

## Scatter Plots

```{r warning=FALSE}
lifeexp %>% 
  ggplot(aes(x = gnppc, y = lexp)) +
  geom_point(color = "steelblue") +
  theme_minimal() +
  labs(x = "GNP Per Capita", y = "Life Expectancy (years)")
```

```{r warning=FALSE}
lifeexp %>% 
  ggplot(aes(x = gnppc, y = lexp)) +
  geom_point(color = "steelblue", size = 10, alpha = 0.7) +
  theme_minimal() +
  labs(x = "GNP Per Capita", y = "Life Expectancy (years)")
```

```{r warning=FALSE}
lifeexp %>% 
  ggplot(aes(x = log(gnppc), y = lexp)) +
  geom_point(color = "steelblue", size = 10, alpha = 0.7) +
  geom_smooth(method = lm, color = "violetred4") +
  theme_minimal() +
  labs(x = "log GNP Per Capita", y = "Life Expectancy (years)")
```

```{r warning=FALSE}
lifeexp %>% 
  ggplot(aes(x = log(gnppc), y = lexp, label = country)) +
  geom_point(color = "white") +
  geom_text(labels = lifeexp$country, size = 3) +
  theme_minimal() +
  labs(x = "log GNP Per Capita", y = "Life Expectancy (years)")
```

